---
title: "The relationship between movie runtime and average user rating"
subtitle: 
author: "Lindsey Coover, Renee Minten, Marlie Snoeijen, Nikki Valen, Fleur van Vugt"
output: pdf_document 
---
```{r setup_packages, include = FALSE, eval = FALSE}
knitr::opts_chunk$set(warning = FALSE)
tinytex:: install_tinytex()

#Include this code chunk if you have not installed tinytex yet (change eval=FALSE TO eval=TRUE)
```

```{r setup_libraries, echo = FALSE, include = FALSE}
knitr::opts_chunk$set(warning = FALSE)
library(tinytex)
library(tidyverse)
library(readr)
```

```{r input_data, warning=FALSE, include=FALSE}
title_basics <- read_tsv("../../data/datasets/title_basics.tsv")
title_ratings <- read_tsv("../../data/datasets/title_ratings.tsv")
data_cleaned <- read_csv("../../gen/data-preparation/temp/data_cleaned.csv")
```

# Introduction

The average length of the ten highest-grossing movies in North America is slightly over two hours (Statista, 2023). This is notable, since survey results among British residents suggest that the ideal length for a movie is under two hours (Dahlgreen, 2015). This prompts the question whether there is a relationship between the runtime of a movie and its success. This research therefore addresses this problem by analyzing data on movie runtimes and their average user ratings. The specific research question addressed in this research is: _What is the relationship between the runtime and the average user rating for movies?_ Results from our analysis may be useful in the movie industry. With knowledge on the effect of the runtime on user rating, producers are able to improve a movie's popularity and therefore, profitability. 

More detailed information for this research project can be found on https://github.com/course-dprep/movie-runtimes-on-average-rating_IMDB. This repository includes a README.md with the project's goal, as well as the instructions and source code to replicate the data preparation and analysis. 

This document is structured as follows. The following section explains the methods that have been used during this research. After, the results of the analysis are presented. Finally, the research question is answered in the conclusion. 

# Data preparation

To prepare the data for analysis, we have operationalized several methods. The steps that we have taken are summarized in the following sub-sections.

## Raw data
 This analysis uses two datasets that are publically available from IMDB. The dataset title_basics contains information on the title of the movie, the year it was released, the genre and the runtime in minutes. The title_ratings dataset consists of the average rating and the number of votes. Summary statistics for the relevant variables in our research are presented below. A more elaborate analysis can be found in the data_exploration.Rmd file on Github. 
 
### Titletype
```{r summary statistics, echo=FALSE, warning=FALSE, include=TRUE}
table(title_basics$titleType)
```

### Run time in minutes 
```{r summary statistics 1, echo=FALSE, warning=FALSE, include=TRUE}
summary(title_basics$runtimeMinutes)
```

### Genres 
```{r summary statistics 2, echo=FALSE, warning=FALSE, include=TRUE}
summary(title_basics$genres)
```

### Average rating
```{r summary statistics 3, echo=FALSE, warning=FALSE, include=TRUE}
summary(title_ratings$averageRating)
```

### Number of votes
```{r summary statistics 4, echo=FALSE, warning=FALSE, include=TRUE}
summary(title_ratings$numVotes)
```

### Start year
```{r summary statistics 5, echo=FALSE, warning=FALSE, include=TRUE}
summary(title_basics$startYear)
```


## Merging process
Subsequently, the columns from the two datasets were merged with a left_join function. This function retains all observations in the title_basics dataset and appends the observations from title_ratings dataset with matching instances. The merged dataset contains 10,194,391 observations.

## Cleaning process
The data cleaning process started with filtering the merged dataset for movies only. Since our research is focused on movies, the observations of other title types (e.g. tvseries, videogames) are irrelevant. After, several variables were renamed and an empty column (endYear) was removed to improve the quality of the dataset. To allow for analysis later in the research, the classes of several variables were converted. Next, the missing observations were removed for the variables _runtimeMinutes_ and _averageRating_. This decision is based on the fact that the research question is mainly based on these two variables. This operation reduced the number of observations to 267,942.
The next part of the cleaning process consisted of removing outliers that could skew the results of the analysis. A review of the box plots for the variables concluded that there are multiple movies with a relatively long run time. This box plot is showed below. Based on this plot, we decided to remove any movies with a run time of more than 300 minutes (5 hours). The merged data was also checked for duplicates, but there were none. Lastly, feature engineering was used to construct two extra variables.  

```{r plots, echo=FALSE, message=FALSE, warning=FALSE}
title_basics$runtimeMinutes <- as.numeric(title_basics$runtimeMinutes)
boxplot(title_basics$runtimeMinutes, ylab = 'runtimeMinutes')
```
More information on the operations used to clean the data can be found in the clean_data.R document on GitHub. 

# Analysis

To estimate the relationship between the runtime of a movie and it's average user rating, a simple linear regression was performed. The analysis consists of two parts.

## Binary linear regression
First, the effect of the run time on the average rating is estimated in a binary regression model. This model only includes the runtime as the independent variable. The output of the linear regression is presented below.
```{r binary_regression_model, message=FALSE, echo=FALSE}

imdb_lm1 <- lm(averageRating ~ runtimeMinutes , data_cleaned)
summary(imdb_lm1)
```
The coefficient for the _runtimeMinutes_ variable is 0.0028217 and is significant at the 5% level. This implies that the run time has a significant positive effect on the run time in minutes of movies. 

## Multiple linear regression
To adjust for the effect of omitted variables in the binary regression model, a multiple regression model with several covariates was estimated as well. These covariates are the number of votes, the year the movie was released and whether the movie is adult-only. As *average* rating is the dependent variable, this is directly correlated to the number of votes and therefore included as a covariate. The start year is included to see if perhaps the quality of movies have gone up throughout the years and therefore the average rating, and adult-only is included to investigate if there is a difference in average rating between adult-only and non-adult-only movies. The results from this regression are visible in the following output. 
```{r multiple_regression_model, message=FALSE, echo=FALSE}
imdb_lm2 <- lm(averageRating ~ runtimeMinutes + numVotes + year + isAdult, data_cleaned)
summary(imdb_lm2)
```
The coefficient for the run time has decreased a little to 0.001963. The effect is still significant at the 5% level. The coefficients for the covariates are also significant at the five percent level.

# Conclusion
We can conclude from our analysis that the run time of a movie has a small, but positive effect on the average rating. This suggests that a longer movie receives a higher average rating. As this effect is relatively small, it is important to take limitations and suggestions for further research into account. Firstly, as the introduction already suggested, there might be cultural differences in the perception of an 'ideal run time'. Further research could focus on average ratings from specific countries, continents or cultures, to see if there is a difference amongst them. Next to that, run time is probably not the most important factor for consumers in rating a certain movie. The narrative of the movie and the acting are factors that play a more important role (Robin Piree, 2023). However, these factors are more difficult to quantify and were not available in the dataset.  

All in all, this analysis shows that there is a small positive relationship between run time in minutes of movies and the average rating. Further research could focus on adding moderators or other independent variables, as well as looking at differences between different countries, continents or cultures. 

# References

Dahlgreen, W. (September 26, 2015). Note to producers: the ideal movie length is under 2 hours. _YouGovUk_, retrieved from https://yougov.co.uk/society/articles/13396-note-producers-perfect-movie-length-under-2-hours?redirect_from=%2Ftopics%2Flifestyle%2Farticles-reports%2F2015%2F09%2F26%2Fnote-producers-perfect-movie-length-under-2-hours

Statista Research Department. (January 5, 2023). Average length of the top 10 highest-grossing movies in the United States and Canada from 1980 to 2021. _Statista_, retrieved from https://www.statista.com/statistics/1292523/lenght-top-movies-us/

Piree, R. (June 7, 2023). What makes a good movie? Retrieved from https://robinpiree.com/blog/what-makes-a-good-movie#:~:text=There%20is%20no%20single%20%22most,heart%20of%20a%20great%20movie. 